---
title: "p8130_hw3_jsg2145"
author: "Jared Garfinkel"
date: "10/25/2019"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(DescTools)
knitr::opts_chunk$set(
	echo = FALSE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

## Problem 2

```{r}
surv_df <- read_csv("./SurvCancer.csv") %>% 
  janitor::clean_names() %>% 
  mutate(organ =
           as.factor(organ),
         organ = 
           fct_reorder(organ, survival, .fun = median))
```

This is a dataset of the survival time of different types of cancer.

This dataset is `r nrow(surv_df)` rows by `r ncol(surv_df)` columns.

There are `r n_distinct(pull(surv_df,organ))` types of organs in the dataset.

```{r}
surv_df %>% 
  count(organ) %>% 
  rename(Organ = organ,
         Count = n) %>% 
  pivot_wider(
    names_from = Organ,
    values_from = Count
  ) %>% 
  knitr::kable(caption = "Count of Subjects.")
```

```{r}
surv_df %>% 
  group_by(organ) %>% 
  summarize(
    Min = min(survival),
    "Qrt 1" = quantile(survival, .25),
    Median = median(survival),
    "Qrt 3" = quantile(survival, .75),
    Max = max(survival)
  ) %>% 
  rename(Organ = organ) %>% 
  knitr::kable()
```

```{r}
surv_df %>% 
  ggplot(aes(x = organ, y = survival)) +
  geom_boxplot() + 
  labs(x = "Organ",
       y = "Survival Time")
```

```{r}
surv_df %>% 
  lm() %>% 
  anova() %>% 
  knitr::kable()
```

$H_0: \bar{X_1} = \bar{X_2} = ... = \bar{X_n}$

$H_1: \bar{X_i} \ne \bar{X_j}, i \ne j$

Reject $H_0$ when $F_{stat} > F_{crit} = F_{k-1,~n-k,~1-\alpha/2}=F_{4,~59,~.995}$ = `r qf(.995, 4, 59)`

Fail to reject $H_0$ when $F_{stat} \le F_{crit}$

$F_{stat} = \frac{SS~Between/(k-1)}{SS~Within/(n-k)} = \frac{2883940}{448274}$ = `r 2883940/448274`

Since $F_{stat} > F_{crit}$, there is sufficient evidence to conclude a significant difference between the 

the sample means (p = .0002295).

```{r}
pairwise.t.test(surv_df$survival, surv_df$organ, p.adj = "bonf", alternative = "two.sided")
```


```{r}
model1 <- surv_df %>% 
  lm() %>% 
  aov()

TukeyHSD(model1, conf.level = .01)
```

```{r}
DunnettTest(surv_df$survival ~ surv_df$organ, control = "Breast")
```

In Bonferroni, Tukey, and Dunnett adjustments a change is made to alpha 

to correct for the family-wise error rate. These results show that in all 

three methods, survival times of patients with breast cancer were significantly 

different at alpha = .01 from stomach cancer, bronchus cancer, and colon cancer.

```{r, fig.width = 10}
surv_df %>% 
  ggplot(aes(x = survival)) +
  geom_histogram(bins = 5) +
  facet_grid(.~organ, scales = "free_x") +
  labs(y = "Count",
       x = "Survival",
       caption = "Frequency of Survival.") +
  scale_y_continuous(breaks = c(2,4,6,8,10))
```

```{r}
surv_df %>% 
  ggplot(aes(sample = survival)) +
  stat_qq() +
  stat_qq_line()
```


```{r}
uhat<-resid(model1)

shapiro.test(uhat)
```

$H_0$: The data are sampled from a population having a normal distribution

$H_1$: The data are sampled from a population not having a normal distribution

An observation of the distribution of data in this dataset shows the data is 

not normally distributed. A qq-plot allows one to visualize the observations compared

to the theoretical normal. The Shapiro-Wilk normality test confirms this, returning

a very low p-value, indicating that the assumption that the data are normal can be 

rejected. One solution to this is to use a non-parametric test that depends on ranked

values rather than on a normality assumption. A Kruskal-Wallis test is acceptable here

because it extends the Wilcoxon rank-sum test to more than two groups.

```{r}
kruskal.test(survival ~ organ, data = surv_df)
```

This test confirms there are significant differences between the survival times 

of patients with different types of cancer at a p-value of 0.01. Compare the Kruskal-Wallis

rank sum test to the parametric test, which had a p-value of .00023. The Kruskal-Wallis test

is more conserviate. With further pairwise testing it is possible to determine 

which pairs of groups are significantly different.

```{r}
pairwise.wilcox.test(surv_df$survival, surv_df$organ, p.adj = "bonf")
```

This table shows that without the normality assumption, none of the groups are 

significantly different with a Bonferroni adjustment.

## Problem 3

```{r}
cats <- c("placebo", "iron", "zinc", "zinc_iron")
n <- c(56, 54, 54, 55)
u <- c(1.1, 1.4, .7, .8)
se <- c(0.2, 0.2, 0.1, 0.1)
epi_df <- tibble(
  cats, n, u, se
)
```

```{r}
epi_df %>%
  rename(Treatment = cats,
         "Sample Size" = n,
         "Average # of Episodes in a Year" = u,
         "Standard Error" = se) %>% 
  knitr::kable()
```

The mean of the groups is a weighted average.

$\mu_{group} = \frac{n_1*\mu_1+n_2*\mu_2}{n_1+n_2}$

```{r}
epi_zinc_df <- tibble(
  cats = c("non_zinc", "zinc"),
  n = c(56+54, 54+55),
  u = c((56*1.1+54*1.4)/(56+54), ((54*.7+55*.8)/(54+55))),
  var = c((((56-1)*2.24)+(54-1)*2.16)/(56+54-2),(((54-1)*.54)+(55-1)*.54)/(54+55-2))
)
```

The effect size is $\mu_{non-zinc}-\mu_{zinc}$ = 1.25-.75 = 0.5 episodes per year

The standard deviation of the groups can be considered to be the standard error

times the square root of the sample size.

$se = \sqrt{\frac{\sigma^2}{n}}$

```{r}
epi_df %>% 
  mutate(
    sd = se*sqrt(n),
    var = sd^2
  ) %>% 
  rename("Treatment" = cats,
         "Sample Size" = n,
         "Average # of Episodes in a Year" = u,
         "Standard Error" = se,
         "Variance" = var,
         "Standard Deviation" = sd) %>% 
  knitr::kable()
```

```{r, F-test, eval = FALSE}
tibble(
  f_crit_nozinc_l = qf(.05, 55, 53),
  f_nozinc = 2.24/2.16,
  f_crit_nozinc_u = qf(.95, 55, 53),
  f_crit_zinc_l = qf(.05, 53, 54),
  f_zinc = .54/.55,
  f_crit_zinc_u = qf(.95, 53, 54)
) %>% 
  knitr::kable()

```

Because the standard deviations are very similar they can be assumed to be the same.

So, the variance of the groups can be calculated using pooled variance.

$s^2 = \frac{(n_1-1)*s_{1}^{2}+(n_2-1)*s_{2}^{2}}{n_1+n_2-2}$

The sample size can be calculated using the following equations for equal allocations:

$n = \frac{(\sigma_{1}^{2}+\sigma_{2}^{2})(z_{1-\alpha/2}+z_{1-\beta})^2}{\Delta^2}$

$z_{1-\alpha/2}$ = `r qnorm(.975)`

$z_{1-\beta}$ = `r qnorm(.8)`

So,

= $\frac{(2.2+0.54)(1.96+.84)^2}{(.5)^2}$

= `r (2.2+.54)*(1.96+.84)^2/(.5)^2` = 85.9 ~ 86

Each group should have 86 subjects to reach this power and significance.

For the sample size of unequal allocations, the following equations apply:

$n_1 = \frac{(\sigma_{1}^{2}+\sigma_{2}^{2}/k)(z_{1-\alpha/2}+z_{1-\beta})^2}{\Delta^2}$

$n_2 = \frac{(k*\sigma_{1}^{2}+\sigma_{2}^{2})(z_{1-\alpha/2}+z_{1-\beta})^2}{\Delta^2}$

So,

$n_1 = \frac{(2.2+0.54/2)(z_{1-\alpha/2}+z_{1-\beta})^2}{(.5)^2}$

= `r (2.2+.54/2)*(1.96+.84)^2/.5^2` = 77.5 ~ 78

$n_2 = \frac{(2*2.2+0.54)(z_{1-\alpha/2}+z_{1-\beta})^2}{(.5)^2}$

= `r (2*2.2+.54)*(1.96+.84)^2/.5^2` = 154.9 ~ 155

The non-zinc group should have 1832 subjects, and the zinc group should have 916 subjects.